Supporting Information for\\ ``A weakly structured stem for human origins in Africa''

### 2022-01-05 supplement.tex
add FST panel plot

The approach from \citet{Ragsdale2020-nz} provides unbiased estimates of the LD
statistics considered here, with smaller sample sizes causing greater
uncertainty in the estimated statistics, but is accounted for by computing
variances/covariances via bootstrap. \sgcomment{I don't think that this is true
if our bootstrap is over genomic regions. In an infinite genome with a tiny
sample size, we would estimate no uncertainty...  Or at least, it is only true
if the Neandertal form a truly randomly mating population...  }

Some statistics, such as $D^2$, require at least two diploid samples to
compute. Since we used a single Neanderthal sample, these statistics for the
Neandertal population were not used in the fit. By contrast, there are
statistics that only require a single sample per population to estimate. These
include cross-population heterozygosity, as well as some statistics involving
more than one population. For example, statistics of the form
$D_{human}(1-2p_{human})(1-2q_{neanderthal})$ require a single Neanderthal
sample and are informative of the Neanderthal demography. These statistics were
included in the fit, but statistics requiring more than one Neanderthal sample
to estimate were removed.

@@

@@

\begin{figure}[ht]
    \centering
    %\includegraphics{figures/supp-FST-panels}
    \caption{
        \textbf{Predicted \bm{$F_{ST}$} between contemporaneous populations over time.}
    }
    \label{fig:supp-FST-panels}
\end{figure}

@@

\begin{figure}[ht]
    \centering
    \includegraphics{figures/supp-csfs-single-origin.pdf}
    \caption{caption}
    \label{fig:supp-csfs-single-origin}
\end{figure}

\begin{figure}[ht]
    \centering
    \includegraphics{figures/supp-csfs-continuous-migration.pdf}
    \caption{caption}
    \label{fig:supp-csfs-continuous-migration}
\end{figure}

\begin{figure}[ht]
    \centering
    \includegraphics{figures/supp-csfs-merger-without-stem-migration.pdf}
    \caption{caption}
    \label{fig:supp-csfs-merger-without-stem-migration}
\end{figure}

\begin{figure}[ht]
    \centering
    \includegraphics{figures/supp-csfs-merger-with-stem-migration.pdf}
    \caption{caption}
    \label{fig:supp-csfs-merger-with-stem-migration}
\end{figure}

>>>>

The approach from \citet{Ragsdale2020-nz} provides unbiased estimates of the LD
statistics considered here, which can be directly compared to expectations for
those same statistics computed from the multi-population Hill-Robertson system.
Estimated statistics for populations with smaller sample sizes have greater
uncertainty, which is largely accounted for by computing variances and
covariances via bootstrap, assuming our samples are unrelated and the
population is randomly mating.

\sgcomment{I don't think that this is true if our bootstrap is
over genomic regions. In an infinite genome with a tiny sample size, we would
estimate no uncertainty...  Or at least, it is only true if the Neandertal form
a truly randomly mating population...  }

Some statistics, such as $\E[D_i D_j]$, require at least two diploid samples
from a population to compute. Since we used a single Neanderthal sample, such
statistics for the Neandertal population were not used in the fit. By contrast,
there are statistics that only require a single sample in a given population to
estimate. These include cross-population nucleotide diversity measures, as well
as some LD statistics involving more than one population. For example,
statistics of the form $D_{human}(1-2p_{human})(1-2q_{neanderthal})$ require a
single Neanderthal sample and are informative of the Neanderthal demography.
These statistics were included in the fit, but statistics requiring more than
one Neanderthal sample to estimate were removed.

@@

$F_{ST}$ can be interpreted as a measure of population structure, given by the
proportion of genetic variance in a multi-population sample that can be
attributed to drift between populations in that sample.

@@

\begin{figure}[ht]
    \centering
    \includegraphics[width=\textwidth]{figures/supp-fsts-all-models.pdf}
    \caption{
        \textbf{Predicted \bm{$F_{ST}$} between contemporaneous populations over time.}
        \aprcomment{Plot ordering, line colors and styles, etc need to be cleaned up}
    }
    \label{fig:supp-FST-panels}
\end{figure}

@@

\begin{figure}[ht]
    \centering
    \includegraphics{figures/supp-csfs-single-origin.pdf}
    \caption{
        \textbf{Conditional SFS compared to single-origin model.}
    }
    \label{fig:supp-csfs-single-origin}
\end{figure}

\begin{figure}[ht]
    \centering
    \includegraphics{figures/supp-csfs-continuous-migration.pdf}
    \caption{
        \textbf{Conditional SFS compared to continuous-migration model.}
    }
    \label{fig:supp-csfs-continuous-migration}
\end{figure}

\begin{figure}[ht]
    \centering
    \includegraphics{figures/supp-csfs-merger-without-stem-migration.pdf}
    \caption{
        \textbf{Conditional SFS compared to merger-without-stem-migration model.}
    }
    \label{fig:supp-csfs-merger-without-stem-migration}
\end{figure}

\begin{figure}[ht]
    \centering
    \includegraphics{figures/supp-csfs-merger-with-stem-migration.pdf}
    \caption{
        \textbf{Conditional SFS compared to merger-with-stem-migration model.}
    }
    \label{fig:supp-csfs-merger-with-stem-migration}
\end{figure}
